MCC Models

### 2019-06-26 papers/full_paper/SI_appendix.Rmd
add prose to numerical example

@@

```{r childcor,  include = T, fig.cap = "Correlation plot for model predictions and data for all models considered in Experiment 7. Coefficients and p-values are based on Pearson correlation statistics. Dots represent condition modes. Error bars represent 95\\% HDIs.", fig.height = 7}

>>>>

### Numerical Example

In the following, we walk through a numerical example for how model predictions were generated for the pragmatics model. The prediction will correspond to the parameter free model, excluding the noise parameter. The numbers are taken from the preference - different speaker - congruent condition. Fig. 1 shows a screenshot from the adult experiment, which the model was designed to capture. In this context, there are three potential referents (two red-ish, one yellow-ish), of two types (red-ish and yellow-ish) on two tables. In principle, the speaker, the frog in this case, can perform one of two utterances: $u_1$, which is turning to the right table (two objects) and making a request or $u_2$, which involves turning to the left table (single object) and making a request. In the experiment, the speaker chose $u_1$ and we therefore model $L_1$'s interpretation of this utterance.

```{r figS1, include = T, fig.align = "center", fig.cap = "Screenshot from adult experiment.", out.width="400px"}
knitr::include_graphics("./figures/SI_setup.png")
```

From a learning perspective, we assume that the listener is learning the mappings between labels and object types instead of object tokens. That is, the listener either thinks the novel word refers to "red-ish objects" or "yellow-ish objects".

The prior distribution over objects (left to right in Fig.1), based on Experiment 2A, for this condition was [`r round(1-mean(adult_ex2_preference_data%>%filter(condition == "different_speaker")%>%pull(correct)),2)`; `r round(1-mean(adult_ex2_preference_data%>%filter(condition == "different_speaker")%>%pull(correct)),2)`; `r round(mean(adult_ex2_preference_data%>%filter(condition == "different_speaker")%>%pull(correct)),2)`]. That is, the yellow object has a higher a priory probability compared to red objects. We start with calculating the informativity of of the utterance for each referent (red or yellow object) based on equation 3.

$$
\begin{aligned}
&Informativity(u_1; r_1) = log(0.64 \cdot 1) = -0.44 \\
&Informativity(u_1; r_2) = log(0.36 \cdot 1) = -1.03
\end{aligned}
$$

with $r_1$ being the yellow object and $r_2$ being the red object. Next we can insert these values into equation 2 to compute the probability that the speaker chooses this utterance to refer to this type of object. To do so, we need the speaker optimality parameter of $\alpha = 2.24$, which was estimated based on Experiment 1.

$$
\begin{aligned}
&P_{S_1}(u_1|r_1) = \frac{exp(2.24 \cdot -0.44)}{exp(2.24 \cdot -0.44)} = 1 \\
&P_{S_1}(u_1|r_2) = \frac{exp(2.24 \cdot -1.03)}{exp(2.24 \cdot -1.03) + exp(2.24 \cdot 0)} = 0.09
\end{aligned}
$$

Here, the informativity for a referent is normalized across all utterances that could be used to refer to this referent. Finally, based on equation 1 we can use these values to compute the probability that the speaker thinks that the speaker is referring to the yellow object ($r_1$) when they produce $r_1$:

$$
\begin{aligned}
&P_{L_1}(r_1,L|u_1) = \frac{1 * 0.64}{1\cdot0.64 + 0.09\cdot0.36 + 0.09\cdot0.36} = 0.91
\end{aligned}
$$

Thus, we normalize across all potential states that the utterance could refer to.

@@

```{r childcor,  include = T, fig.cap = "Correlation plot for model predictions and data for all models considered in Experiment 7. Dots represent condition modes. Error bars represent 95\\% HDIs.", fig.height = 7}
